\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}
\section{History}
Here an attempt will be made to very briefly highlight some of the ideas, features and syntax of a number of languages. The classification is quite arbitrary as most languages support multiple paradigms.
\section{Assembly languages}
RISC >< CISC
\subsection{M32}
\subsection{Z80}
\subsection{(M)MIX}
\subsection{x86}
\subsection{ARM}

\section{Procedural languages}
\subsection{FORTRAN 77}
\input{lang/fortran77}
\subsection{Fortran 95}
\input{lang/fortran95}
\subsection{C}
\subsection{Pascal}
Of course, in stupid languages like Pascal, where labels cannot be
descriptive, GOTOs can be bad. But that's not the fault of the goto,
that's the braindamage of the language designer.


\subsection{SNOBOL}
\subsubsection{History}
Development of SNOBOL (StriNg Oriented and symBOlic Language) began in 1962 at AT\&T Bell Labs by Ralph Griswold, Ivan Polonsky and David Farber. The goal was to develop a string processing language for formula manipulation and graph analysis. The most successful version was SNOBOL4, developed in the 1970s.

SPITBOL (Speedy Implementation of SNOBOL) is a compiled implementation of the SNOBOL4.
\subsubsection{Concepts}
\begin{itemize}
\item Pattern matching language based upon BNF Grammars.
\item Patterns are a first-class data type.
\end{itemize}
\subsection{Go}

\section{UML}
\input{lang/UML}

\section{Object-oriented languages}
\subsection{Ada}
\subsection{C++}
\subsection{Smalltalk}
\subsection{Java}
\subsection{Scala}
\input{lang/scala}
\subsection{Eiffel}
Design by contract.

\section{Interpreted languages}
These are general-purpose programming languages that typically support multiple paradigms.
\subsection{Python}
special * syntax

 PEP 3107 -- Function Annotations
 PEP 484 -- Type Hints
\subsubsection{Implementations}
CPython, PyPy.
\subsection{Javascript}
\subsection{Lua}
\subsection{Perl}
\subsection{R}
\input{lang/r}

\section{Functional languages}
\subsection{Functional programming paradigm}
pure state, 
\subsection{LISP}
\subsection{Scheme}
Continuations?
\subsection{ML}
\subsection{Haskell}
\input{lang/haskell}

\section{Logic programming languages}
\subsection{Planner}
Inspired smalltalk and prolog
\subsection{Prolog}
\input{lang/prolog}

\section{Numerical computing}
\subsection{MATLAB}
\subsection{GNU Octave}
\subsection{Maxima}
\subsection{Julia}

\section{Shell scripting}
\subsection{Bash}
\subsection{DOS}

\section{System and circuit design}
\subsection{SPICE}
\subsection{Verilog}
\subsection{Chisel}
\subsection{FIRRTL}
\subsection{LabVIEW and G}

\section{Database querying}
\subsection{SQL}
\section{Data-interchange}
\subsection{XML}
\subsection{JSON}

\addtocontents{toc}{\protect\setcounter{tocdepth}{5}}