\chapter{Introduction}
Space is obviously quite important for physics. Finding a mathematical model for space presents some challenges. Many branches of mathematics have tried to model essential characteristics of space in different ways. This has lead to many different mathematical meanings, one of which we have already seen: the vector space.

\begin{displayquote}
SPACE. The word came into English—from Old French from Latin—around 1300. The OED entry distinguishes many meanings. In one sense (under heading 6b) it has room as a synonym. This word derives from the Old English and is related to the modern German Raum. Under heading 17 the OED defines “a space” as “an instance of any of various mathematical concepts, usually regarded as a set of points having some specified structure.” Among the quotations is a nice one from 1932: “The word ‘space’ has gradually acquired a mathematical significance so broad that it is virtually equivalent to the word ‘class’, as used in logic.” (M. H. Stone Linear Transformations in Hilbert Space p. 1.) The space age was well under way by 1914 when Hausdorff’s Grundzüge der Mengenlehre (Fundamentals of Set Theory) gave axioms for a METRIC SPACE (metrischer Raum) and for a TOPOLOGICAL SPACE (topologischer Raum).\footnote{From \textit{Earliest Known Uses of Some of the Words of Mathematics}, \url{http://jeff560.tripod.com/s.html}}
\end{displayquote}

The main branch of mathematics that is relevant for the modeling of space is obviously geometry. Just like the word ``space'', the label geometry is applied to many parts of mathematical reasoning.

\section{Erlangen programm}
In 1872 Felix Klein proposed his \udef{Erlangen programm} (named after the University Erlangen-Nürnberg, where Klein worked) which attempted to classify different geometries based on symmetries. In particular it allowed geometry to be viewed as the study of properties of figures that remain invariant under a certain group of transformations. So for example, in Euclidean geometry may be viewed as the study of properties that remain invariant under isometry transformations (these can roughly be viewed as the translations and rotations). Two shapes are called \udef{congruent} if there is an isometry that maps one onto the other. If we take a rectangle, its corners will always be $90\si{\degree}$ angles no matter how we rotate or translate it. Angles are in general invariant under such transformations. Other invariants include
\begin{itemize}
\item Distances;
\item Areas;
\item Volumes;
\item Whether lines are parallel, or not;
\item Whether points are or other shapes, or not;
\item Whether points are collinear, or not;
\end{itemize}

Another important aspect to the Erlangen program is its hierarchical nature: if we take a larger group of transformations, then fewer aspects will invariants of all the transformations. Conversely, if we restrict the group of transformations, more aspects will be invariants. For example, the isometry transformations are part of a larger group of transformations called affine transformations. In particular they are the affine transformations that preserve distance. Of the bulleted list of invariants above, the last three are affine invariants, but the first three are \textbf{not}.

In this way projective geometry may be seen as the underlying, unifying frame. Restricting it a bit we get affine geometry; restricting it a bit more we get Euclidean geometry.

\section{Analytic–synthetic distinction}
TODO

\chapter{Euclidean and related geometry}
\section{Axiomatic (or synthetic) Euclidean geometry}
Historically it was easy: geometry was the study of shapes, either on a flat plane or in space (or what was somewhat naively thought to be space). By the third century BC Euclid had published his \textit{Elements}. In it he gave gave the postulates of what is now called Euclidean geometry. Euclid held his postulates to be self-evidently applicable the the actual, physical space in the real world.

\subsection{Euclid's \textit{Elements}}
There are 13 books\footnote{It must be remembered that in ancient times book binding technology was not as advanced and works were published in smaller subunits called \textit{books}. Each book was only a few dozen pages long, in today's pages, but published on scrolls.} in the \textit{Elements}.
\begin{itemize}
\item Books I to IV and VI discuss planar geometry.
\begin{itemize}
\item[\textbf{Book I}] lays out the fundamentals of planar geometry involving straight lines.
\item[\textbf{Book II}] contains propositions to do with squares and rectangles. This has a strong link with geometric algebra due to the link with the squaring of a number.  
\item[\textbf{Book III}] lays out the fundamentals of planar geometry involving circles.
\item[\textbf{Book IV}] deals with the intersection of circles and rectilinear figures.
\item[\textbf{Book VI}] discusses similar figures.
\end{itemize}
\item  Books V and VII-X deal with number theory, with numbers treated geometrically as lengths of line segments or areas of regions.
\item Books XI-XIII concern solid geometry.
\item There are apocryphal books XIV and XV, probably written by Hypsicles and Isidore of Miletus, respectively. These are not usually included.
\end{itemize}

What follows is the first part of book I\footnote{Translation by TODO ref}, which contains some definitions and Euclid's postulates.

\begin{displayquote}
{\LARGE \centering BOOK I \par}
{\centering DEFINITIONS\par}
\begin{enumerate}
\item A \textbf{point} is that which has no part.
\item A \textbf{line} is breadthless length.
\item The extremities of a line are points.
\item A \textbf{straight line} is a line which lies evenly with the
points on itself.
\item A \textbf{surface} is that which has length and breadth only.
\item The extremities of a surface are lines.
\item A \textbf{plane surface} is a surface which lies evenly with
the straight lines on itself.
\item A \textbf{plane angle} is the inclination to one another of
two lines in a plane which meet one another and do not lie in
a straight line.
\item And when the lines containing the angle are straight,
the angle is called \textbf{rectilineal}.
\item When a straight line set up on a straight line makes
the adjacent angles equal to one another, each of the equal
angles is \textbf{right}, and the straight line standing on the other is
called a \textbf{perpendicular} to that on which it stands.
\item An \textbf{obtuse angle} is an angle greater than a right
angle.
\item An \text{acute angle} is an angle less than a right angle.
\item A \textbf{boundary} is that which is an extremity of anything.
\item A \textbf{figure} is that which is contained by any boundary
or boundaries.
\item A \textbf{circle} is a plane figure contained by one line such
that all the straight lines falling upon it from one point among
those lying within the figure are equal to one another;
\item And the point is called the \textbf{centre} of the circle.
\item A \textbf{diameter} of the circle is any straight line drawn
through the centre and terminated in both directions by the
circumference of the circle, and such a straight line also
bisects the circle.
\item A \textbf{semicircle} is the figure contained by the diameter
and the circumference cut off by it. And the centre of the
semicircle is the same as that of the circle.
\item \textbf{Rectilineal figures} are those which are contained
by straight lines, \textbf{trilateral} figures being those contained by
three, \textbf{quadrilateral} those contained by four, and \textbf{multilateral}
those contained by more than four straight lines.
\item Of trilateral figures, an \textbf{equilateral triangle} is that
which has its three sides equal, an \textbf{isosceles triangle} that
which has two of its sides alone equal, and a \textbf{scalene
triangle} that which has its three sides unequal.
\item Further, of trilateral figures, a \textbf{right-angled triangle}
is that which has a right angle, an \textbf{obtuse-angled
triangle} that which has an obtuse angle, and an \textbf{acute-angled
triangle} that which has its three angles acute.
\item Of quadrilateral figures, a \textbf{square} is that which is
both equilateral and right-angled; an \textbf{oblong} that which is
right-angled but not equilateral; a \textbf{rhombus} that which is
equilateral but not right-angled; and a \textbf{rhomboid} that which
has its opposite sides and angles equal to one another but is
neither equilateral nor right-angled. And let quadrilaterals
other than these be called \textbf{trapezia}.
\item \textbf{Parallel} straight lines are straight lines which,
being in the same plane and being produced indefinitely in
both directions, do not meet one another in either direction.
\end{enumerate}
{\centering POSTULATES\par}
Let the following be postulated :
\begin{enumerate}
\item To draw a straight line from any point to any point.
\item To produce a finite straight line continuously in a
straight line.
\item To describe a circle with any centre and distance.
\item That all right angles are equal to one another.
\item That, if a straight line falling on two straight lines
make the interior angles on the same side less than two right
angles, the two straight lines, if produced indefinitely, meet
on that side on which are the angles less than the two right
angles.
\end{enumerate}
{\centering COMMON NOTIONS\par}
\begin{enumerate}
\item Things which are equal to the same thing are also
equal to one another.
\item If equals be added to equals, the wholes are equal.
\item If equals be subtracted from equals, the remainders
are equal.
\item Things which coincide with one another are equal to
one another.
\item The whole is greater than the part.
\end{enumerate}
\end{displayquote}

The rest of book I contains propositions. The other books contain definitions and propositions.

When Euclid uses ``equal'', he means equal in magnitude. This is effectively what is now called congruence. This equals is obviously an equivalence relation (the transitivity and reflexivity are asserted as ``common notions'' 1. and 4.). Common notions 2. and 3. assert that the operations of addition and subtraction are compatible with the equality relation.

There is some debate as to whether the common notions were (in part or at all) included by Euclid\footnote{See ref TODO for much insightful commentary}. 

In the postulates, Euclid only explicitly asserts the \textit{existence} of the constructed objects, in his reasoning \textit{uniqueness} is implicitly assumed. 

\subsubsection{The fifth postulate.} Reading through the postulates, it may be obvious that the fifth is of quite a different nature than the rest. It looks like is should be a proposition and for centuries mathematicians tried, unsuccessfully, to prove it as such.

Euclid himself, it seems, mistrusted the postulate and proved the first 28 propositions without using it.

Many equivalent formulations of the fifth postulate exist, the most famous probably being Playfair's axiom:

\begin{displayquote}
In a plane, through a point not on a given straight line, at most one line can be drawn that never meets the given line.
\end{displayquote}

In fact there is always exactly one. It can be shown (and in fact follows from proposition 27) that even in absolute geometry we can always find at least one parallel line.

Only in the beginning of the $19^\text{th}$ century did mathematicians start what Euclidean geometry would look like without the fifth postulate. Leaving out this postulate, one obtains what is known as \udef{absolute geometry}.

If the fifth postulate were provable as a theorem, absolute geometry would be the same as Euclidean geometry. It turns out that this is not the case and including the negation of the fifth postulate leads to a consistent set of axioms, describing a \udef{non-Euclidean geometry}.

\subsection{Other axiomatic systems}
Euclid's axioms do not actually provide the complete, logical foundation he thought they did. Many authors have offered their own sets of axioms that meet modern standards or rigour. Moritz Pasch was the first to accomplish this task in 1882.

\subsubsection{David Hilbert} proposed a set of axioms in 1899 that did not depart too greatly in spirit from Euclid's, but was complete. The result was an axiom system constructed with six primitive notions (point, line, plane, betweenness, congruence and containment) and twenty axioms devided into 5 classes (incidence, order, congruence, parallels and continuity).

\subsubsection{George Birkhoff} created a set of axioms\footnote{TODO ref.} for planar geometry that uses real numbers in 1932. Leveraging the mathematics of real numbers, only four axioms were needed. Birkhoff's reason for introducing them was that they may be readily verified in the real world using a ruler and a protractor. Their simplicity also allowed them to be used in high-school books.

The primitive terms are:
\begin{enumerate}[(a)]
\item \textbf{Points}, designated by $A,B,C, \ldots$
\item Particular sets of points called \textbf{lines}, designated by $l,m, \ldots$
\item The symmetric relation \textbf{distance} between any two points $A,B$, designated by $d(A,B)$
\item The \textbf{angle} determined by three ordered points $A,O,B \;(A\neq O, B\neq O)$, designated $\angle AOB$, is a relation between two lines, $AO$ and $BO$.
\end{enumerate}
Then the axioms may be stated as follows:
\begin{enumerate}
\item[\textbf{Postulate I}] \textit{Postulate of Line Measure}. The points $A, B, \ldots$ of any line can be put into 1:1 correspondence with the real numbers $x$ so that $|x_B - x_A| = d(A, B)$ for all points $A$ and $B$. 
\item[\textbf{Postulate II}] \textit{Point-Line Postulate}. There is one and only one straight line, $l$, that contains any two given distinct points $P$ and $Q$.

\begin{definition}
Before continuing with the rest of the postulates, we must interject with a few definitions.
\begin{itemize}
\item For any three points $A,B,C$ on the same line (i.e.\ \udef{collinear points}), point $B$ is said to be \udef{between} $A$ and $C$ if
\[ d(A,C) = d(A,B) + d(B,C) \]
\item A \udef{line segment} $AC$ is the set of points on the line through $A$ and $C$ that are between $A$ and $C$. Equivalently, it is the set of points $P$ such that
\[ x_A \leq x_P \leq x_C \quad \text{or} \quad x_C \leq x_P \leq x_A \]
\item A \udef{ray} or \udef{half-line} $l'$ with \udef{end-point} $O$ is defined by two distinct points $O,A$ on line $l$ as the set of all points $A'$ on $l$ such that $O$ is not between $A$ and $A'$. TODO fig.
\item A \udef{broken line} $ABC\ldots KL$ consists of a collection of segments $AB, BC, CD, \ldots, KL$. The points $A,B, \ldots, L$ are the \udef{vertices} of the broken line.
\item If the initial point $A$ and the terminal point $L$ coincide, the broken line is called a \udef{polygon}.
\item A polygon with three distinct vertices is called a \udef{triangle}.
\end{itemize}
\end{definition}

\item[\textbf{Postulate III}] \textit{Postulate of Angle Measure}. The rays {$l, m, n, \ldots$} through any point $O$ can be put into 1:1 correspondence with the real numbers $a \mod 2\pi$ so that if $A$ and $B$ are points (not equal to $O$) of $l$ and $m$, respectively, the difference $(a_m - a_l) \mod 2\pi$ of the numbers associated with the lines $l$ and $m$ is the angle $\angle AOB$. Furthermore, if the point $B$ on $m$ varies continuously in a line $r$ not containing the vertex $O$, the number $a_m$ varies continuously also. 
\item[\textbf{Postulate IV}] \textit{Postulate of Similarity}\footnote{This postulate rules out non-Euclidean geometries}. If in two triangles $\triangle ABC$ and $\triangle A'B'C'$  and for some constant $k > 0$,
\[d(A', B' ) = k\cdot d(A, B), \quad d(A', C' ) = k\cdot d(A, C) \quad \text{and} \quad \angle B'A'C'  = \pm \angle BAC, \]
then
\[d(B', C' ) = k\cdot d(B, C), \quad \angle  C'B'A'  = \pm \angle CBA, \quad \text{and}\quad  \angle A'C'B'  = \pm \angle ACB. \]
\end{enumerate}

\begin{definition}
Some more definitions:
\begin{itemize}
\item As a consequence of postulate II, two distinct lines have either one point in common, or none. In the first case they are said to \udef{intersect} in their common point; in the second case, they are said to be \udef{parallel}.
\item Two figures are called \udef{similar} if all corresponding distances are in proportion and all corresponding angles are equal or all negatives of each other.
\item Two figures are called \udef{congruent} if they are similar with a ratio of proportionality equal to one.
\end{itemize}
\end{definition}

\paragraph{Modifications to allow for higher dimensions}

\section{Analytic Euclidean geometry}

In the $17^\text{th}$ century René Descartes and Pierre de Fermat departed from this purely axiomatic (or \udef{synthetic}) approach and introduced the \udef{analytic} approach, explicitly using a coordinate system. This was an important reason for developing linear algebra (as lines and planes can be represented by linear equations) and one of the reasons we talk about vector \textit{spaces}. In fact $n$-dimensional Euclidean space can be modeled using an $n$-dimensional vector space with the standard inner product (that supplies the notions of distance and angle).

For the rest of the geometries mentioned here, we will focus on the analytic side of things, as that approach is more useful for the practicing physicist.

\subsection{Introducing the model}
All this talk of axioms may be frustrating for an engineer of physicist who just wants to be able to calculate. We now introduce a model (in fact a class of models) that can easily be used to calculate with. It is of course important to verify that these models do in fact describe Euclidean geometries. We will do that by verifying that they satisfy Birkhoff's postulates.

The models are based on real vector spaces equipped with the dot product, together with definitions for the primitive terms point, line, distance and angle.

TODO:

position and displacement vector

Zero

dimension

$V$

Using bold $\vec{v}$ for vectors

Taking a look at the axiom for angle measurement, it is obvious that we need a way to determine the angle between two vectors where the angle can be anything from $0$ to $2\pi$. The problem with this is that the dot product only gives us the cosine of the angle $\cos{\theta}$. Inverting that, we get a number between $0$ and $\pi$. In other words, always the smallest angle between two vectors (TODO fig). What we need to do in order to get an number between $0$ and $2\pi$ is fix an \textbf{orientation}. To do that we need a basis (TODO need?). We can then multiply the result of the arc cosine by $-1$ and take the angle mod $2\pi$ if the orientation of the vectors is negative.

\begin{note}
\begin{enumerate}
\item[\textbf{Point}] A point $A$ is modeled by a vector $\vec{v}_A \in V$.
\item[\textbf{Line}] A line is any set of vectors of the following form
\[ l = \{ \vec{v}_A + \lambda(\vec{v}_B - \vec{v}_A) \;|\; \lambda \in \R\} \]
where $\vec{v}_A$ and $\vec{v}_B$ are distinct vectors in $V$. Conventionally this is denoted
\[ l \; \leftrightarrow \; \vec{v}_A + \lambda(\vec{v}_B - \vec{v}_A). \]
\item[\textbf{Distance}] The distance between points $\vec{v}_A$ and $\vec{v}_B$ is given by
\[ d(A,B) = \lVert \vec{v}_B - \vec{v}_A \lVert = \sqrt{(\vec{v}_B - \vec{v}_A)\cdot (\vec{v}_B - \vec{v}_A)} \]
\item[\textbf{Angle}] The angle $\angle AOB$ is given by
\[ \angle AOB = \pm\cos^{-1}\left(\frac{(\vec{v}_A- \vec{v}_O)\cdot (\vec{v}_B- \vec{v}_O)}{\lVert \vec{v}_A - \vec{v}_O\lVert\cdot\lVert \vec{v}_B- \vec{v}_O\lVert}\right) \]
\end{enumerate}
The angle is negative if $\left((\vec{v}_A- \vec{v}_O), (\vec{v}_B- \vec{v}_O)\right)$ has a negative orientation.
\end{note}

\subsection{Compatibility with Birkhoff's postulates}
TODO: lots of figures

It turns out it's easiest to consider the postulates in a different order, so that is what we will do.

\begin{enumerate}
\item[Postulate II] This proof contains two parts:
\begin{enumerate}
\item \textit{Existence}: for any two points a straight line can be found that contains both points.
\item \textit{Uniqueness}: only one such line can be found. We need to show that any such line we can construct is equivalent.
\end{enumerate}
The proof is as follows:
\begin{enumerate}
\item Existence is easy. Take two arbitrary points $P,Q$. Consider the line
\[ l \;\leftrightarrow\; \vec{v}_P + \lambda (\vec{v}_Q - \vec{v}_P) \]
setting $\lambda = 0$ we see that the line contains $\vec{v}_P$; setting $\lambda = 1$ we see that the line contains $\vec{v}_Q$. So this line is a good line.
\item Say we have another straight line $m$ such that $m$ contains $\vec{v}_P$ and $\vec{v}_Q$. The line $m$ can be written as
\[ m \;\leftrightarrow\; \vec{v}_A + \mu (\vec{v}_B - \vec{v}_A)  \]
for some $\vec{v}_A$ and $\vec{v}_B$. We must show that $m = l$. We split this into two parts: $l \subset m$ and $m \subset l$.
\begin{enumerate}
\item[$\boxed{l \subset m}$] Because $\vec{v}_P, \vec{v}_Q \in m$ there must exist $\mu_P, \mu_Q \in \R$ such that
\begin{equation}
\begin{cases}
\vec{v}_A + \mu_P(\vec{v}_B - \vec{v}_A) = \vec{v}_P \\
\vec{v}_A + \mu_Q(\vec{v}_B - \vec{v}_A) = \vec{v}_Q.
\end{cases} \label{vPvQ}
\end{equation}
These expressions for $\vec{v}_P$ and $\vec{v}_Q$ can be filled in in the expression for the line $l$:
\begin{align}
l \;\leftrightarrow\; &\vec{v}_P + \lambda (\vec{v}_Q - \vec{v}_P) \\
&\vec{v}_A + \mu_P(\vec{v}_B - \vec{v}_A) + \lambda ((\vec{v}_A + \mu_Q(\vec{v}_B - \vec{v}_A)) - (\vec{v}_A + \mu_P(\vec{v}_B - \vec{v}_A))) \\
&\vec{v}_A + \mu_P(\vec{v}_B - \vec{v}_A) + \lambda (\mu_Q(\vec{v}_B - \vec{v}_A) - \mu_P(\vec{v}_B - \vec{v}_A)) \\
&\vec{v}_A + [\mu_P + \lambda(\mu_Q - \mu_P)](\vec{v}_B - \vec{v}_A).
\end{align}
For every $\lambda \in \R$, the expression $(\mu_P + \lambda(\mu_Q - \mu_P))$ is a real number and thus a value $\mu$ can take. This means that every point of $l$ is also a point of $m$ and thus $l \subset m$.
\item[$\boxed{m \subset l}$] Because $\vec{v}_P$ and $\vec{v}_Q$ are distinct (and thus $\mu_P \neq \mu_Q$), equations (\ref{vPvQ}) can be inverted to obtain
\[ \begin{cases}
\vec{v}_A = \vec{v}_P + \frac{-\mu_P}{\mu_P - \mu_Q}(\vec{v}_Q - \vec{v}_P) \\
\vec{v}_B = \vec{v}_P + \frac{\mu_P-1}{\mu_P - \mu_Q}(\vec{v}_Q - \vec{v}_P)
\end{cases} \]
With a very similar line of reasoning, we can see that $m \subset l$.
\end{enumerate}
This concludes the proof.
\end{enumerate}
\item[Postulate I] Many such bijections can be found, each corresponding with a different placement and orientation of the ruler used. Assume that the points $A,B,C,D$ are on the line $l$ and are distinct. We elect to place the beginning of our ruler at point $C$and consider the half of the line on which $D$ lies as being in the positive direction. Consider the function
\[ x: l\to \R: A \mapsto x_A = \pm\lVert \vec{v}_A - \vec{v}_C \lVert \]
where the expression for $x_A$ is positive if $C$ is not between $A$ and $D$. We now need to prove two things
\begin{enumerate}
\item The proposed mapping $x$ is a 1:1 correspondence (i.e.\ a bijection) and
\item The distance $d(A,B) = \lVert \vec{v}_B - \vec{v}_A\lVert$ is equal to $|x_B - x_A|$.
\end{enumerate}
We proceed as follows:
\begin{enumerate}
\item We first introduce the special unit vector 
\[ \hat{v}_l \equiv \frac{\vec{v}_D - \vec{v}_C}{\lVert \vec{v}_D - \vec{v}_C\lVert} \]
Because $D$ and $C$ lie on the line and taking into account the second postulate, we see that we can write the line $l$ as
\[ l \;\leftrightarrow\; \vec{v}_C + \lambda \hat{v}_l. \]
\begin{itemize}
\item Now to prove surjectivity, we need to prove that for any real number $y$. There is a point $\vec{v}_E$ on the line such that $x_E = y$. Take an arbitrary real number $y$. The claim is now that the relevant point is given by $\vec{v}_E = \vec{v}_C + y \hat{v}_l$. Indeed
\[ x_E = \pm\lVert (\vec{v}_C + y \hat{v}_l) - \vec{v}_C \lVert = \pm\lVert y \hat{v}_l \lVert = \pm |y| \lVert \hat{v}_l \lVert = \pm |y|. \]
Now this is positive if $\vec{v}_E$ is on the $D$ side of $\vec{C}$, which is exactly the case if $y$ is positive, so $x_E = y$.
\item Injectivity states that if we have two distinct points $A,B\in l$, then $x_A \neq x_B$. To prove this, write $A$ and $B$ as
\[ \begin{cases}
\vec{v}_A = \vec{v}_C + y_A \hat{v}_l \\
\vec{v}_B = \vec{v}_C + y_B \hat{v}_l
\end{cases} \]
which must necessarily be possible for some $y_A, y_B \in \R$ with $y_A\neq y_B$. Reasoning as before we obtain
\[ \begin{cases}
x_A = y_A \\ x_B = y_B.
\end{cases} \]
Thus $x_A \neq x_B$.
\end{itemize}
\item As before we write
\[ \begin{cases}
\vec{v}_A = \vec{v}_C + y_A \hat{v}_l \\
\vec{v}_B = \vec{v}_C + y_B \hat{v}_l
\end{cases} \]
Consequently
\begin{align}
d(A,B) &= \lVert \vec{v}_B - \vec{v}_A\lVert \\
&= \lVert (\vec{v}_C + y_B \hat{v}_l) - (\vec{v}_C + y_A \hat{v}_l)\lVert \\
&= \lVert y_B \hat{v}_l - y_A \hat{v}_l\lVert \\
&= |y_B - y_A|\cdot\lVert \hat{v}_l\lVert \\
&= |y_B - y_A|
\end{align}
and
\begin{align}
|x_B - x_A| = |y_B - y_A|.
\end{align}
\end{enumerate}
This concludes the proof.
\item[Postulate III] The proof that our model satisfies this axiom is similar to the last one. We will again propose a mapping that we will show to be a bijection with the requisite properties. We choose a ray $n$ to act as our reference. For any ray $l$ with a point $A$ on it, we define the unit vector along the ray $\hat{e}_l$ as
\[ \hat{e}_l = \frac{\vec{v}_A - \vec{v}_O}{\lVert \vec{v}_A - \vec{v}_O \lVert}. \]
As a consequence of the second postulate these unit vectors are unique.
Then we define the function $a$ on the rays through $O$ as follows:
\[ a_l = \pm \cos^{-1}(\hat{e}_l \cdot \hat{e}_n) \mod 2\pi \]
where the minus sign appears if $(\hat{e}_l, \hat{e}_n)$ has a negative orientation.

We also define $\hat{e}_t$ as the unique unit vector that makes $(\hat{e}_n, \hat{e}_t)$ a positively oriented orthonormal basis (TODO ?).

Now we need to show that
\begin{enumerate}
\item The mapping $a$ is a bijection.
\item If $A$ and $B$ are points (not equal to $O$) of rays $l$ and $m$ through $O$, then 
\[ \angle AOB = (a_m - a_l)\mod 2\pi \]
\item If $B$ varies continuously, then $a_m$ varies continuously also.
\end{enumerate}
We proceed as follows:
\begin{enumerate}
\item We first prove injectivity and then surjectivity.
\begin{itemize}
\item To prove injectivity we take two rays $l$ and $m$. Assuming that $a_l = a_m$, we need to show that $l=m$. Clearly $a_l = a_m$ implies that
\[ \hat{e}_l \cdot \hat{e}_n = \hat{e}_m \cdot \hat{e}_n \]
and that $(\hat{e}_l, \hat{e}_n)$ and $(\hat{e}_m, \hat{e}_n)$ have the same orientation.
The unit vector $\hat{e}_l$ has the following orthonormal decomposition:
\[ \hat{e}_l = (\hat{e}_l \cdot \hat{e}_n)\hat{e}_n + (\hat{e}_l \cdot \hat{e}_t)\hat{e}_t \]
Using the fact that it is a unit vector, we get
\[ \hat{e}_l^2 = (\hat{e}_l \cdot \hat{e}_n)^2 + (\hat{e}_l \cdot \hat{e}_t)^2 = 1 \]
Thus $\hat{e}_l \cdot \hat{e}_t = \pm\sqrt{1- (\hat{e}_l \cdot \hat{e}_n)^2}$. This shows that $\hat{e}_l$ is one of two vectors. For one of those the orientation of $(\hat{e}_l, \hat{e}_n)$ is positive, for it is negative (TODO: show). Same for $\hat{e}_m$. Thus because the orientation of $(\hat{e}_l, \hat{e}_n)$ and $(\hat{e}_m, \hat{e}_n)$ is the same, $\hat{e}_l$ and $\hat{e}_m$ are the same vector. Then considering the rays through $\vec{v}_O$ and $\vec{v}_O + \hat{e}_l$, and through $\vec{v}_O$ and $\vec{v}_O + \hat{e}_m$, postulate II gives $l=m$ and the sought-after injectivity.
\item For surjectivity we must find a ray $l$ for every angle in $[0,2\pi[$ such that $a_l$ equals that angle. Take an arbitrary angle $\theta \in [0,2\pi[$. TODO
\end{itemize}
\item TODO
\item TODO
\end{enumerate}
This concludes the proof.
\item[Postulate IV] TODO
\end{enumerate}

\subsection{Towards categoricity}
TODO $\mathbb{E}$
\subsection{Spatial analytic geometry}

\section{Projective geometry}

Also in the $17^\text{th}$ mathematicians were trying to see what happened if certain axioms or concepts were left out of Euclids axiomatic system.
In particular Girard Desargues started the systematic study of projective geometry, which does not have any concept of distance or parallel lines. The original motivation for this was an attempt to understand perspective.

There are several axiomatisations of projective geometry (such as those by Whitehead, Coxeter, Hilbert \& Cohn-Vossen and Greenberg). We will not be considering those.

From an analytic point of view, the $n$-dimensional projective space over an arbitrary field $K$ can be constructed as follows:
\begin{itemize}
\item Take an $n$-dimensional vector space $V$ over the field $K$. We define $K_0$ and $V_0$ as resp. the sets $K$ and $V$ without the neutral element for the addition, $0$. i.e.\
\[ V_0 \equiv V\setminus \{0\} \qquad K_0 \equiv K\setminus \{0\} \]
\item We define the equivalence relation $\sim$ on $V_0$:
\[ v \sim w \;\Leftrightarrow\; \exists \lambda \in K_0: v = \lambda w. \]
It should be clear that this is indeed an equivalence relation.
\item We define $[v]$ as the equivalence class that contains $v$. Explicitly this is given by
\[ [v] = \{ \lambda v \; |\; \lambda \in K_0 \}. \]
This gives a partition of $V_0$ (i.e.\ $[v] = [w] \Leftrightarrow v\sim w$ and $v \nsim w \Rightarrow [v] \cap [w] = \emptyset$).
\item The \udef{projective space} $P(V)$ associated to $V$ can now be defined as the set of equivalence classes. In other words, it is a quotient space:
\[ P(V) \, = \,V_0 / \sim\; =  \,\{ [v] \;|\; v \in V_0 \}  \]
Each equivalence class is called a \textit{point} of the projective space. It may be strange to call a set a point (TODO point about models and isomorphism)
\item In the construction above we have collapsed whole lines (containing e.g\ the points $\lambda v$ for all $\lambda \in K$) into single points. It therefore makes sense to \textit{define} the dimension of $P(V)$ to be one less than the dimension of $V$, if $V$ has a finite dimension that is.
\end{itemize}
In order to get the full projective geometry, according to the Erlanger program, we now also need a group of transformations. For the construction proposed above, the projective transformations $\phi: P(V) \to P(W)$ can be constructed as follows.
\begin{itemize}
\item Take the vector space $\GL(V,W)$ of the bijective linear maps from $V$ to $W$. We would like to define the projective transformations as the transformations of the form
\[ P(V) \to P(W): [v] \mapsto [f(v)] \]
where $f$ is an element of $\GL(V,W)$. It is not immediately clear that this well defined. The problem is that $[v]$ is a set of which $v$ is only one element. If we take a different element of $u \in [v]$, how do we know that $f(u) \in [f(v)]$? In other words how do we know that projective transformation does not depend on the (arbitrarily chosen) representative $v$ of the projective point $[v]$?
\item Luckily we can use the result that for all $v\in V_0$, $[f(v)] = [g(v)]$ if and only if $[f] = [g]$. The notation $[f]$ makes sense because $\GL(V,W)$ is itself a vector space.
\end{itemize}


\section{Affine geometry}
Tangent space and bundle

\section{Back to Euclidean geometry}

\section{Non-Euclidean geometry}
The development of non-Euclidean geometries (with different sets of axioms) was another exciting enrichment of the field. These axiomatic systems roughly describe shapes in space that is not flat. In 2D this translates to doing geometry on surface that is not flat, like a sphere. From an analytic viewpoint, vector spaces (embodying linearity) are obviously no longer 

Tangent bundle!!

\input{differentialGeometry}

\input{LieGroups}

\input{riemannianManifolds}




\chapter{Algebraic geometry}
\chapter{Geometric topology}
\section{Vector bundles}
\subsection{Definition}
\subsection{Operations on vector bundles}