\documentclass{report}
\input{../preamble}

\makeatletter
\@fileswtrue
\makeatother

\title{Some of the Big Ideas in Mathematics}
\author{Joseph Cunningham}
\date{}

\addbibresource{mathematics.bib}

\begin{document}
\maketitle
\tableofcontents

TODO: Bertrand paradox (probability); Routh-Hurwitz; Marden's theorem, Sylvester-Gallai; \url{https://mathoverflow.net/questions/28997/does-anyone-know-an-intuitive-proof-of-the-birkhoff-ergodic-theorem}

\url{https://grossack.site/}

\chapter{Preface}
The text is a collection of interesting results. If a proof is not supplied, it it probably fairly trivial, but it may also be an oversight. This text starts with set theory and attempts to develop as much mathematics as possible from there. Only a basic knowledge of logic is assumed.

In particular the following logical symbols will be used and not introduced:
\begin{center}
\begin{tabular}{ c c }
Symbol & Meaning \\
 \hline
$\land$ & logical and \\
$\lor$ & logical or \\
$\neg$ & logical not \\
$\forall$ & universal quantifier \\
$\exists$ & existential quantifier \\
$\exists!$ & uniqueness quantifier \\
$(\,), [\,]$ & parentheses \\
$=$ & equality \\
$\neq$ & non-equality
\end{tabular}
\end{center}

One important note for proving things: if we say $\exists x: P(x)$, we can take this $x$ and continue the proof using it.

\section{I am a formalist}
I think it is very hard to argue that many of the mathematical objects discussed here \textit{actually} exist. You will not find a real number with infinite precision floating around in the universe. Most real numbers, i.e.\ the ones that do not have a simple representation or construction procedure, cannot even be thought of.

Still, they seem to be very useful to model many things in this world, so we want a way to study them. Since I do not really believe the physical existence, what am I studying? Some abstract thought? A meme (in the original meaning)? If so, how do I know that I am talking about the `same' numbers as everyone else? This is essentially the second horm of Benacerraf's dilemma.

Also, what does it mean to be wrong about a mathematical object? Say I had convinced myself that all numbers larger than one hundred are even and then somebody tells me I am wrong. Of course I ask for proof and they say ``Look, 101 is larger than 100 and not even''. Why should I accept this? I have two lines of reasoning, leading to opposite conclusions. Which is more trustworthy? And why?
In this case, clearly the counterexample is generally considered more trustworthy, but in principle I could have rejected that 101 was even a proper number. Since they live in my head, I can do what I want. This feels very wrong, because we have so many physical analogues of natural numbers all around us. It is not too difficult to assemble 101 marbles and use that as a justification for the existence of 101.

For natural numbers that are small enough we have good justification and we have a robust way to communicate observations about them. We can be fairly sure we are actually thinking of the same things. For large numbers this becomes a problem. Consider $9^{10^{11^{12}}}$ and suppose I want to know what the middle digit of the decimal expansion is. There is no way I can write down the full decimal expansion. If I have an argument to say it is 1 and my friend has one to say that it is 2, on what grounds should we believe one over the other? The situation becomes even more complicated when considering real numbers like $\pi+e+\sqrt{2}$.

Of course, in practice, mathematicians have a way to sort good arguments from bad ones. They use formal logic. If me and my friend go to a mathematician and ask who is correct, the mathematician will presumably ask us the explain our proofs step by step and will at the end declare there to be an error in one (or both) of them.

The mathematician is judging the steps by comparing them to a finite set of finite rules. It is exactly this abstract machinery that is used to share the idea of, say, real numbers. The abstract machine is finite and we can easily build physical models of it (by manipulating symbols on a piece of paper or by programming a computer). Given these physical models, we can validate the ideas we send and receive. The idea of the real numbers can then be shared by sharing the idea of the machine together with a (finite) recipe for constructing the real numbers in it.

We can trust the abstract machine in a way that we cannot trust very large of very precise numbers. In this way I am a finitist. But I am a finitist when it comes to the supporting machine, not when it comes to the numbers themselves. I see this as having all the advantages of finitism with fewer of the downsides. For example, a (strict) finitist cannot say much about all natural numbers. They cannot say that all natural numbers are even or odd. Rather they must say that all natural numbers less that a million are even or odd. Or a billion. Whatever they can verify. This can be quite cumbersome. With the support of the machine, I can easily say all natural numbers are even or odd.

Another big advantage is that the machine can be adapted to suit many different situations. Given a nice calculation, we do not need to ponder the metaphysical reality of what has been written. Rather, it is usually relatively easy to make the necessary definitions accessible to the machine such that we can prove this theorem, say about lattice-valued groups or some other relatively esoteric mathematical `object'.

I value arguments and calculations. I will invent any context necessary that naturally supports an interesting argument. I have the feeling many mathematicians feel the same way.

TODO: this also means mathematics can yield good models for science: just ask the machine and it will give you an answer. The computation may use esoteric mathematical `objects', but the output should be testable.

\part{Set Theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{setTheory}
\input{setTheory2}

\part{Formal Languages and Term Rewriting}
\setcounter{chapter}{0} % Reset chapter counter
\input{formalLanguages}

\part{Order Theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{orderTheory}

\part{Operator Structures}
\setcounter{chapter}{0} % Reset chapter counter
\input{operatorStructures}

\part{Category Theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{categoryTheory}

\part{Model Theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{modelTheory}

\part{Algebra}
\setcounter{chapter}{0} % Reset chapter counter
\input{algebra}

\part{Discrete Mathematics}
\setcounter{chapter}{0} % Reset chapter counter
\chapter{Summation}
\[ a\sum_{i=1}^n x_i = \sum_{i=1}^n ax_i \]
\[ \sum_{i=1}^n\sum_{j=1}^m x_{ij} = \sum_{j=1}^m\sum_{i=1}^n x_{ij} \]

$\delta_{ij}$

Introduce $a:b$ for sequence. (bounds inclusive)

$-$ means everything.

\begin{proposition} \label{productOfSum}
Let $R$ be a ring, $n\in \N$ and $\seq{a_i},\seq{b_i}\in R^n$. Then
\[ \prod_{i=0}^{n-1}(a_i + b_i) = \sum_{\seq{s_i}\in \{0,1\}^n}\prod_{i=0}^{n-1}\bigg(k\mapsto \begin{cases}
a_i & (k=0) \\ b_i & (k=1)
\end{cases}\bigg)(s_i). \]
\end{proposition}

\begin{proposition} \label{sumAsDecimation}
Let $G$ be an Abelian group, $f:\N\to G$ a function and $M,N \geq 1 \in \N$. Then
\[ \sum_{k=0}^{M\cdot N-1}f(k) = \sum_{k_0 = 0}^{M-1}\sum_{k_1 = 0}^{N-1} f(k_1\cdot M + k_0). \]
\end{proposition}
\begin{proof}
We prove this by induction on $N$. For the base case $N=1$, we have
\begin{align*}
\sum_{k=0}^{M\cdot N-1}f(k) &= \sum_{k=0}^{M-1}f(k) \\
&= \sum_{k_0=0}^{M-1}f(k_0) \\
&= \sum_{k_0=0}^{M-1}\sum_{k_1 = 0}^{0}f(k_1 \cdot M + k_0) \\
&= \sum_{k_0=0}^{M-1}\sum_{k_1 = 0}^{N-1}f(k_1 \cdot M + k_0).
\end{align*}
For the induction step, assume the induction hypothesis, i.e.\ that the statement holds for $N-1$. Then
\begin{align*}
\sum_{k=0}^{M\cdot N-1}f(k) &= \sum_{k=0}^{M\cdot (N-1)-1 + M}f(k) \\
&= \sum_{k=0}^{M\cdot (N-1)-1}f(k) + \sum_{k=M\cdot (N-1)}^{M\cdot (N-1)-1 + M}f(k) \\
&= \sum_{k_0 = 0}^{M-1}\sum_{k_1 = 0}^{N-2} f(k_1\cdot M + k_0) + \sum_{k_0=0}^{M-1}f(M\cdot (N-1) + k_0) \\
&= \sum_{k_0 = 0}^{M-1}\Big(\sum_{k_1 = 0}^{N-2} f(k_1\cdot M + k_0) + f(M\cdot (N-1) + k_0)\Big) \\
&= \sum_{k_0 = 0}^{M-1}\sum_{k_1 = 0}^{N-1} f(k_1\cdot M + k_0).
\end{align*}
By induction we conclude that the statement holds for all $M,N \geq 1$.
\end{proof}

\input{combinatorics}

\part{Elements of Mathematics}
\setcounter{chapter}{0} % Reset chapter counter
\chapter{Elements of Euclidean geometry}
\input{euclideanGeometry}

\part{Convergence and Topology}
\setcounter{chapter}{0} % Reset chapter counter
\input{topology}

\part{Number Systems}
\setcounter{chapter}{0} % Reset chapter counter
\input{numbers}

\part{Linear Algebra}
\setcounter{chapter}{0} % Reset chapter counter
\input{linearAlgebra/linearAlgebra}

\part{Convergence Vector Spaces}
\setcounter{chapter}{0} % Reset chapter counter
\input{convergenceVectorSpaces}

\part{Functional Analysis}
\setcounter{chapter}{0} % Reset chapter counter
\input{functionalAnalysis}

\part{Analysis}
\setcounter{chapter}{0} % Reset chapter counters
\input{analysis}

\part{Hilbert Spaces and $C^*$ Algebras}
\setcounter{chapter}{0} % Reset chapter counter
\input{hilbertSpacesCstarAlgebras}

\part{Operator Equations}
\setcounter{chapter}{0} % Reset chapter counters
\input{diffEq}

\part{Probability Theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{probabilityTheory}

\part{Statistics}
\setcounter{chapter}{0} % Reset chapter counters
\input{stats}

\part{Geometry}
\setcounter{chapter}{0} % Reset chapter counter
\input{geometry}

\part{Number Theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{numberTheory}

\part{$K$-theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{Ktheory}


\part{Applied Mathematics}
\setcounter{chapter}{0} % Reset chapter counters
\chapter{Optimisation}
\input{optimisation}

\chapter{Curves and surfaces in Euclidean space}
\input{curves}
\chapter{Vector and tensor calculus}
\input{vectorCalculus}




\appendix

\chapter{Symbols}


\listoftheorems[title={List of named results}, swapnumber,ignoreall,onlynamed]

%\printindex[definition]


\chapter{Bibliography}

\url{https://en.wikipedia.org/wiki/Philosophy_of_mathematics}

Categories for the working mathematician, Saunders Mac Lane

Mathematical Foundations of Game Theory
An Introduction to Decision Theory ?

Philosophy of Mathematics (Princeton Foundations of Contemporary Philosophy)
\url{https://plato.stanford.edu/entries/philosophy-mathematics/#Cat}
An Introduction to the Philosophy of Mathematics (Cambridge Introductions to Philosophy)
The Oxford Handbook of Philosophy of Mathematics and Logic, Stewart Shapiro
BELIEVING THE AXIOMS. PENELOPE MADDY
Philosophy of Mathematics, Jeremy Avigad
Thinking about mathematics, Shapiro


\begin{itemize}
\item Logicaboek, Batens
\item A friendly introduction to mathematical logic, Christopher C. Leary
\item A course in mathematical logic, John Bell and Moshé Machover
\item Mathematical Logic, Lightstone
\item Handbook of Mathematical Logic, Jon Barwise ed.
\item \url{http://www2.hawaii.edu/~robertop/Courses/TMP/7_Peano_Axioms.pdf}
\item Axiom of choice, Jech
\item \url{https://arxiv.org/abs/0904.4205v1}
\item \url{http://philsci-archive.pitt.edu/2703/1/reconstruction2.pdf}
\item Raymond Smullyan on Self Reference
\item Anticipatory Systems; Philosophical, Mathematical, and Methodological Foundations
\item A Shorter model theory Hodges
\item Marker's Model Theory; An Introduction
\item Equivalence: an attempt at a history of the idea
\item A book of Set Theory, Charles C. Pinter
\item Enderton's "Elements of Set Theory
\item I. V. Volovich, ``Number theory as the ultimate physical theory'', Preprint No. TH 4781/87, CERN, Geneva, (1987)
\item Weyl, Philosophy of Mathematics and Natural Science
\item Heath, Thomas L. (1956). The Thirteen Books of Euclid's Elements (2nd ed. [Facsimile. Original publication: Cambridge University Press, 1925] ed.). New York: Dover Publications. In 3 vols.: vol. 1 ISBN 0-486-60088-2, vol. 2 ISBN 0-486-60089-0, vol. 3 ISBN 0-486-60090-4. Heath's authoritative translation of Euclid's Elements, plus his extensive historical research and detailed commentary throughout the text.
\item Birkhoff, George David (1932), "A Set of Postulates for Plane Geometry (Based on Scale and Protractors)", Annals of Mathematics, 33: 329–345, doi:10.2307/1968336
\item Birkhoff's Axioms for Space Geometry
Roland Brossard
The American Mathematical Monthly
Vol. 71, No. 6 (Jun. - Jul., 1964), pp. 593-606 (14 pages)
\item Backgrounds Of Arithmetic And Geometry: An Introduction
By Branzei Dan, Miron Radu
\item \url{http://lya.fciencias.unam.mx/gfgf/ga20101/material/analyticgeometry_chap1.pdf}
\item Topology, James Munkres
\item Counterexamples in topology
\item Real analysis in reverse: \url{https://arxiv.org/pdf/1204.4483.pdf}
\item A garden of integrals, Frank E. Burk
\item A textbook on ordinary differential equations, Shair and Ambrosetti
\item Partial differential equations, Lawrence C. Evans
\item Partial differential equations in action, Sandro Salsa
\item Wikipedia: real numbers,
\item Categories for the Working Mathematician, Saunders Mac Lane
\item Category Theory for the Sciences (The MIT Press): David I. Spivak
\item \url{https://www.math.mcgill.ca/triples/Barr-Wells-ctcs.pdf}
\item Abraham Robinson (1996):Non-Standard Analysis, revised edition. Princeton University Press, Princeton,New Jersey.
\item Lectures on the Hyperreals
\item The way of the Infinitesimal, Fabrizio Genovese
\item A guide to distribution theory and fourier transforms, Robert S. Strichartz
\item Hall, Brian C. (2003). \textit{Lie Groups, Lie Algebras, and Representations}. New York: Springer.
\item Machì, Antonio. (2012). \textit{Groups : an introduction to ideas and methods of the theory of groups}. Milan: Springer.
\item Brian G. Wybourne, Classical groups for physicists
\item Varadarajan, Lie groups
\item Nastase, Classical field theory
\item Introduction to Differentiable Manifolds, Serge Lang
\item A Comprehensive Introduction to Differential Geometry, Michael Spivak
\item W. Schläg - Complex Analysis and Riemann Surfaces
\item L.W. Tu - An Introduction to Manifolds
\item (Lee)
\item Blanché, Axiomatics
\item Kunen, Set theory, An introduction to independence proofs
\item Feldman, Epistemology
\item BonJour, Epistemology
\item Doets, Basic model theory
\item Hodges shorter model theory
\item Moschovakis, Notes on set theory
\item Theory of formal systems, Smullyan
\item How  to  Prove  It:  A  Structured  Approach
\item \url{https://www.drmaciver.com/2015/12/you-can-tell-about-zorns-lemma/}
\item Istrăţescu, Inner product structures
\item Axler, Linear Algebra Done Right
\item Kreyszig - Introductory Functional Analysis with Applications
\item Conway, A Course in Functional Analysis
\item Helemskii - Lectures and exercises on functional analysis
\item Steven Roman - Advanced Linear Algebra
\item Narici, Beckenstein - Topological Vector Spaces
\item Schmidt, relational mathematics
\item \url{https://www.math.ksu.edu/~nagy/real-an/}
\item Jordan, Smith - Mathematical techniques
\item Dunford N., Schwartz J. - Linear operators
\item Fillmore - A user's guide to operator algebras.
\item Blackadar - Operator Algebras.
\item M.  Rørdam,  F.  Larsen,  N.  Laustsen - An  introduction  to  K-theory  for $C^*$-algebras.
\item Dana P. Williams - Crossed products of C-star algebras
\item Pedersen, Gert Kjaergård - C-star-algebras and their automorphism groups
\item Spin geometry - Lawson
\item Notes on Category Theory - Paolo Perrone
\item Introduction to Tensor Products of Banach Spaces - Ryan
\item Postmodern Analysis - Jost
\item Harmonic Analysis of Operators on Hilbert Space
\item Sheaves in Geometry and Logic
\item Introduction to set theory - Monk
\item Infinite Dimensional Analysis: A Hitchhiker’s Guide
\item Handbook of mathematics - Thierry Vialar
\item Probability: A Graduate Course - Allan Gut
\item Probability: Theory and Examples - Rick Durrett
\item A second course in linear algebra - Garcia, Horn
\item Matrix Analysis - Horn, Johnson
\item Markov processes, semigroups and generators - Kolokoltsov V.N.
\item Topics in Linear and Nonlinear Functional Analysis - Gerald Tesch
\item Calculus on Normed Vector Spaces - Coleman
\item Differential Calculus in Topological Linear Spaces - Yamamuro
\item Galois Connections and Applications
\item Term Rewriting and All That - Franz Baader, Tobias Nipkow
\item Term Rewriting Systems - Terese
\item Ordinary Differential equations - Hartman
\item Convergence foundations of topology - Dolecki, Mynard
\item Techniques of Functional Analysis for Differential and Integral Equations - Sacks
\item A Hilbert Space Problem Book - Halmos
\item Introduction to Operator Theory in Riesz Spaces - Zaanen
\item \url{https://www.math.uwaterloo.ca/~snburris/htdocs/UALG/univ-algebra2012.pdf}
\item Lattices and Ordered Algebraic Structures - T.S. Blyth
\item \url{https://www.mat.univie.ac.at/~gerald/ftp/book-fa/fa.pdf}
\item A Course on Topological Vector Spaces - Voigt
\item Pre-Riesz spaces - Kalauch, van Gaans
\item Convergence Structures and Applications to Functional Analysis - Beattie, Butzmann
\item Continuous Convergence on C(X) - Binz
\item Introduction to Spectral Theory - Hislop, Sigal
\item A Course in Topological Combinatorics - de Longueville
\item \url{https://people.bath.ac.uk/mw2319/ma30252/lecturenotes.html}
\item Frames and Locales, Topology without points - Picado, Pultr
\item \url{https://arxiv.org/pdf/0907.5356.pdf}
\item Clifford Algebras. An Introduction. - Garling
\item An Introduction to Clifford Algebras and Spinors - Vaz, da Rocha
\item Topological vector spaces - Robertson, Robertson
\item Principles of Harmonic Analysis - Deitmar, Echterhoff
\item One-Parameter Semigroups for Linear Evolution Equations - Engel, Nagel
\item Equations of Evolution - Tanabe
\item \url{https://link.springer.com/article/10.1007/s11225-012-9465-0}
\item \url{https://www.bioinf.uni-leipzig.de/~studla/Publications/PREPRINTS/01-pfs-007-subl1.pdf}
\item Fourier Analysis - Körner
\item Fourier Analysis - Folland
\item Classical Fourier Analysis - Grafakos
\item Modern Fourier Analysis - Grafakos
\item Measure Theory - Bogachev
\item Measure Theory - Cohn
\item Measure Theory - Tao
\item Measure and Integration Theory - Bauer
\item Lectures on Analysis on Metric Spaces - Heinonen
\item Geometric Measure Theory - Federer
\item Geometric Measure Theory, A Beginner's Guide - Morgan
\item Measure Theory and Fine Properties of Functions - Evans, Gariepy
\item Axioms for infinite matroids
\item Graph Theory - Wilson
\item Intro to convergence theory with filters: 
url{https://www.bioinf.uni-leipzig.de/~studla/Publications/PREPRINTS/01-pfs-007-subl1.pdf}
\end{itemize}

\printbibliography


\end{document}
