\documentclass{report}
\input{../preamble}
\zexternaldocument*[m-]{../Mathematics/TheBigIdeas_Mathematics}

\title{Some of the Big Ideas in Physics}
\author{Joseph Cunningham}
\date{}

\begin{document}
\maketitle
\tableofcontents

\chapter{Preface}
This text is meant to be a succinct and self contained overview of the topics I have come across during my physics studies. The aim is to give an overview of the most important ideas and theories in physics into a relatively short text. It is also my aim to provide a solid theoretical and conceptual basis for these theories.  




 that is more or less accessible even to people with no background in physics or mathematics. This will almost certainly prove impossible, but I will give it a shot.

It is my view that mathematical notation provides an invaluable tool to support the physical reasoning in a concise way. The notation is often however quite a large barrier for people who have not studied the subject. In an attempt to make this text accessible to everyone I will try to explain all of the mathematics assuming no previous knowledge whatsoever.

It must also be obvious that my explanations and derivations cannot be completely rigorous and fit in a text the size of this one. All I will attempt is to guide the reader through some of the reasoning processes that gave us modern physics, trying to make everything seem plausible along the way. The ultimate goal is to give the reader enough insight so that they could work out the rigorous details on their own.

I doubt very much I will be able to achieve this goal.
I will almost certainly err on the side of terseness due to time constraints, but I would be very grateful if anyone reading this text (especially anyone without a mathematical background) would give me feedback on which parts are understandable and which are not. It is not always obvious as the author how clear one's explanations are.

A major motivation for writing this text is to put into words, as clearly as possible, the logical structure of physics. By that I mean present a framework that allows one to trace back the ``why'' as far as possible; until, that is, we reach the frontiers of mathematical abstraction. I would love to tackle an overview of some of the key concepts in mathematics, but that is most definitely a future project.

Because I am not an expert, I need to rely on people who are. I have therefore borrowed shamelessly from several textbooks, as referenced in the bibliography, and Wikipedia.

TODO! Integrate the following:

In the pages that follow there will be a lot of mathematics; many strange symbols, complex formulae and the like. They all represent ideas, ways of formulating and solving problems to do with physics. I will attempt to formulate here very briefly some of the guiding principles that have been useful for me to understand the subject matter (insofar as I actually understand it, that is) and I will try to expand on these ideas in the body of the text.

Firstly there is the question of how much one needs to know to adequately understand a topic. I feel the professor's answer to this is slightly different from mine. I would generally like slightly more mathematical background. I have added some material to that effect, but it still remains a very cursory introduction to the topic of Lie groups, Lie algebras and representations. My guiding principle when adding the material has been the question of ``why''. Why introduce all these concepts? Why do the calculations we are doing? I have tried to make the story more or less a coherent one, to make the definitions seem not too unnatural and to make the theorems at least plausible. I do however sometimes not observe strict mathematical rigour in favour of a more intuitive approach, in the hopes that the intuition enables the reader to work out the mathematical details if necessary.

Another thing I would like to touch one briefly here is the difference between a definition and a notion in mathematics. I will introduce several definitions for Lie groups and Lie algebras. Which, while quite different, all refer to the same idea or notion of this concept we call Lie groups and algebras. Hopefully the different definitions and the different angles at which the topic is approached will serve to strengthen the intuition of anybody foolish enough to read this text.

\begin{comment}
Finally two general points about why we do what we do in this text. Firstly it is a general goal of mathematics to generalize things and strip away anything that is not strictly necessary. Take for example the notion of a space. We know how to work with normal 3D space. We can measure distances and angles, we can look at neighbourhoods of points etc. Now the mathematician will ask 
- set of point with stuff layered on in order to approximate reality
- mathematical modus operandi (generalisation) -> new physics (cfr. Dirac etc.)
\end{comment}

TODO Carnappian reconstruction; axioms:
In discussing axiomatic systems several properties are often focused on:[6]

    The axioms of an axiomatic system are said to be consistent if no logical contradiction can be derived from them. Except in the simplest systems, consistency is a difficult property to establish in an axiomatic system. On the other hand, if a model exists for the axiomatic system, then any contradiction derivable in the system is also derivable in the model, and the axiomatic system is as consistent as any system in which the model belongs. This property (having a model) is referred to as relative consistency or model consistency.
    An axiom is called independent if it can not be proved or disproved from the other axioms of the axiomatic system. An axiomatic system is said to be independent if each of its axioms is independent. If a true statement is a logical consequence of an axiomatic system, then it will be a true statement in every model of that system. To prove that an axiom is independent of the remaining axioms of the system, it is sufficient to find two models of the remaining axioms, for which the axiom is a true statement in one and a false statement in the other. Independence is not always a desirable property from a pedagogical viewpoint.
    An axiomatic system is called complete if every statement expressible in the terms of the system is either provable or has a provable negation. Another way to state this is that no independent statement can be added to a complete axiomatic system which is consistent with axioms of that system.
    An axiomatic system is categorical if any two models of the system are isomorphic (essentially, there is only one model for the system). A categorical system is necessarily complete, but completeness does not imply categoricity. In some situations categoricity is not a desirable property, since categorical axiomatic systems can not be generalized. For instance, the value of the axiomatic system for group theory is that it is not categorical, so proving a result in group theory means that the result is valid in all the different models for group theory and one doesn't have to reprove the result in each of the non-isomorphic models.

TODO: model theory!!

\chapter{Introduction: why the big ideas?}
TODO + counter examples

\chapter{What is physics?}
Physics is of course a science. Generally the goal of science is to gain an understanding and make predictions about how the world works. This understanding will not just appear in our minds unprovoked and reason alone will not provide all the answers. Consequently we must interact with the real world in order to get information about it. Experiments are such interactions.

Having done an experiment, we have information about the world. A priori this information is not very useful. It says something about the world when the experiment took place. That world is no longer. We must hypothesise that the world is not completely random and thus that it does follow some underlying rules. Unfortunately we cannot directly verify any underlying rule because we can only do so many experiments and for any set of experimental results there are many rules that can explain them. For example, if we drop a ball a hundred times and it always falls down to the ground, this may indicate that it will always fall down. It may however be the case that the hundred and first time it falls upwards. We can easily prove this wrong however by dropping it another time and seeing that it falls down. We can still not be sure however that it will always fall down. This leads us to the important insight that a theory can be proven wrong, but never right. In fact a theory is \textit{only really worth considering} if it makes predictions we can try to prove wrong. In that case we call it \udef{falsifiable}.

These ruminations lead us to view the scientific process as roughly the following: 
\begin{enumerate}
\item First we \textbf{observe} some phenomenon.
\item This inspires us to \textbf{hypothesise} an explanation (often based on and extending theories that have previously been hypothesised and tested).
\item We then \textbf{make predictions} based on our newly hypothesised theory.
\item Finally we \textbf{test} these predictions. After enough predictions have been tested we accept that the hypothesis is probably true.
\end{enumerate}

Having talked a bit about science in general, we now turn to physics in particular. While it's mostly convention which parts of scientific inquiry fall under the label physics, we can generally say that physics is in the business of developing theories and principles that are true for everything in the world. So for example whether you are dropping a ball or a cat, physics provides laws of motion that explain both trajectories. This is in contrast to e.g. biology that can only say something about the behaviour of the cat and what they would do in such a situation.

\section{What is a formula? Why mathematics?}
We need a way to share our ideas about the world. Debate is necessary to further scientific knowledge. No one person can understand all of physics on their own; even if this was possible, that person would probably still need to write things down to order their own thoughts. We physiologically do not have the brain capacity to actively hold all physics concepts in our mind at the same time.

With mathematics a system has been developed that helps us write down our reasoning process in a clear and concise manner. It can be seen as a kind of shorthand that allows us to write down our reasoning in steps so we don't need to keep everything constantly in mind.

A formula is the mathematical formulation of some fact we believe to have uncovered about the universe. It is a crystallisation of our knowledge of the world formulated using mathematics.

In this way formulae can be seen almost as an illustration of a concept. In fact illustrations can provide much of the same illumination as formulae and can even serve to provide additional understanding if the standard notation is too unwieldy. Judicious use of illustrations in this document will hopefully aid in understanding.

\begin{example}
As an illustration of the utility of formulae, we will consider the following problem:

Farmer John has 200 metres of fencing that he can use to fence off an area of one of his fields. He would like to know what the biggest possible surface area is he can fence off, if the fenced off area has to be a rectangle.

We will use some mathematical notation and procedures without much justification here. They will be clarified later; this example will partially help motivate the future discussion. 

As will often be the case, the first step is to give names to things. So, for instance, we might call the length of the field $l$ and the width $w$. The main advantage of this is, when considering complex lines of reasoning, we can take in and process these single letter symbols much faster that always having to reference ``the length of the field''. Thus we can formulate the main constraint of the problem quite simply and very compactly as
\[ 2l + 2w = \SI{200}{m} \]
Where we write $2l$ and $2w$ because the perimeter of a rectangle 2 times the length plus 2 times the width. This means that the width also equals 200 metres minus 2 times the length, all divided by two
\[w = \frac{\SI{200}{m} - 2l}{2} = \SI{100}{m}-l\]

TODO: image x2

In order to get the surface area (which we call $S$) of the field, we multiply the length by the width
\[S = l\cdot w = (\SI{100}{m})\cdot l - l^2 \]
Now the power of the abstract notation and reasoning really gets to shine: We are in effect trying to maximise the value of $S$ and if we had properly introduced the mathematical machinery of calculus we would know that we can do that by setting the derivative of $S$ to zero.
\[ \SI{100}{m} - 2l = 0 \qquad \text{from which follows that} \qquad l = \SI{50}{m} \]
For now we will find a different way to reach the same conclusion, because we have not yet developed the requisite calculus.

There are undoubtedly many ways to solve this problem. Here we will use an approach that is in some sense variational in nature (in that we will look at an expression for the surface area around the maximum). In fact we will prove that regardless of the actual length of the fenching, for a problem like this a square is the best shape. A proof might go as follows:

For a given length of fencing, a square can always be constructed. This square has a particular length that we call $a$ for convenience and a surface area $a^2$. If we now look at other rectangles with the same circumference, they will have a side that is longer (we can say its length is $a+x$ for some \emph{positive} $x$) and a side that is shorter. The shorter side must be shorter by the same amount that the longer side is longer by (in order for the total length of fencing to stay constant). So the shorter side has length $a-x$. See figure ref TODO We can now just work out the surface area:
\[ S = (a+x)\cdot (a-x) = a^2 - x^2 \]
This means that for any nonzero $x$ the surface area is smaller than that of the square and thus the square is the best shape.

A fortiori we conclude that in our particular case $l=\SI{50}{m}$ is optimal as this gives us a square.

\end{example}

This example also exemplifies the general aesthetic we will be going for: Mostly text punctuated with formulae and calculations to provide clarity and set out facts for future reference.

\part{Some experimental notes}
\setcounter{chapter}{0} % Reset chapter counter
\chapter{Units}
TODO flow
Now the (imperial) system of measurements used in the example above is not so very useful for more complex an precise measurements.
A first obvious problem is that not all barleycorns have the same length. This can of course be easily remedied by choosing one particular barleycorn as the standard.
If many people are supposed to use this system of measurement, they cannot all have access to this special barleycorn.
One way to get around this is by letting people copy the length of the barleycorn, in effect making rulers based on this barleycorn. 

Another solution is to describe the unit of measurement in such a way that anybody can make there own ruler. This idea appeared in the seventeenth century and was known as the universal measure or \textgreek{μέτρον καθολικόν} (métron katholikón) in Greek. (Hence the name metre).
One proposal for such a system was the seconds pendulum, which is a pendulum with a period of exactly two seconds. As we will see, pendulums with the same period have the same length if the effect of gravity stays constant, independent of the attached weight.
Unfortunately gravity is not equally strong everywhere (and the second was not undisputed as unit of time at the time) and thus it was decided that the new measure should be one ten-millionth of the distance from the North Pole to the Equator, measured along the meridian through Paris. This required the distance to be carefully measured, which was not an easy task at the time. A platinum bar, the mètre des Archives, was manufactured based on the measurements to serve as the prototype for all other metres. In the following years new surveys revealed that the mètre des Archives was not exactly correct, but for practical reasons it was decided in 1867 that the official length of the should be the length of the mètre des Archives. This meant we were back to being stuck with a definition of our units based on an arbitrary object, except now it was a platinum-iridium bar instead of a barleycorn. It was only in the twentieth century that a true universal definition of the metre was adopted (first using the wavelength of light emitted by krypton-86 and later as the length traveled by light in one 299792458th of a second).

While the project of the metre initially failed to provide a universal measure, it did introduce another useful innovation: it was a decimal system.
In previous systems if the lengths you were measuring were too big or too small to be easily measured in your unit of choice, you would pick another unit that was often a quite arbitrary multiple of the first unit.
So for example you might measure the length of people in feet, but the length of voyages in miles (which is 5280 feet). With the metric system all units derived from the metre differ by multiplication of a power of ten. Thus, e.g., a kilometre (km) is a thousand (which is $10^3$) metres. A more extensive list can be found in table \ref{metricPrefixes}. This made it much easier to quantify very big and very small things, because no matter the scale, there was a logical system of units to work with.

\begin{table}[h]
  \centering
  $\begin{array}{c|c|c | r}
    \text{Prefix name} & \text{Prefix symbol} & \text{Power of Ten} & \text{Decimal factor}\\
	\hline
	\text{yotta} & Y & 10^{24} & 1 000 000 000 000 000 000 000 000 \\
	\text{zetta} & Z & 10^{21} & 1 000 000 000 000 000 000 000 \\
	\text{exa} & E & 10^{18} &  1 000 000 000 000 000 000\\
	\text{peta} & P & 10^{15} & 1 000 000 000 000 000 \\
	\text{tera} & T & 10^{12} & 1 000 000 000 000 \\
	\text{giga} & G & 10^{9} & 1 000 000 000 \\
	\text{mega} & M & 10^{6} & 1 000 000 \\
	\text{kilo} & k & 10^{3} & 1 000 \\
	\text{hecto} & h & 10^2 & 100 \\
	\text{deca} & da & 10^1 & 10 \\
	& & 10^0 & 1 \\
	\text{deci} & d & 10^{-1} & 0.1 \\
	\text{centi} & c & 10^{-2} & 0.01 \\
	\text{milli} & m & 10^{-3} & 0.001 \\
	\text{micro} & \mu & 10^{-6} & 0.000 001 \\
	\text{nano} & n & 10^{-9} & 0.000 000 001 \\
	\text{pico} & p & 10^{-12} & 0.000 000 000 001 \\
	\text{femto} & f & 10^{-15} & 0.000 000 000 000 001 \\
	\text{atto} & a & 10^{-18} & 0.000 000 000 000 000 001 \\
	\text{zepto} & z & 10^{-21} & 0.000 000 000 000 000 000 001 \\
	\text{yocto} & y & 10^{-24} & 0.000 000 000 000 000 000 000 001 \\
  \end{array}$
  \caption{Prefixes for the International System of Units}
  \label{metricPrefixes}
\end{table}

When writing these units down, we often do not want to write out the full unit. Every SI unit has an abbreviation, formed by concatenating the prefix symbol with the abbreviation of the unit, thus we can write pm instead of picometres or ms instead of milliseconds.

\section{Base units}

So far we have mainly talked about units of length, with a brief mention of units of time in passing. Of course we will want to measure things other than just length and time, like temperature, mass, speed, electric current, electric charge, force, pressure etc. It would seem that we need to define units for all these quantities, preferably with universal definitions independent of and physical reference object. Luckily we do not need universal definitions for all these units. For example speed can be expressed as a length traveled per unit of time, thus we can use the unit meters per second (m/s). It turns out we can express many units in terms of other units so that we only need universal definitions for 7 units, the so-called SI base units. These units are listed in table \ref{baseUnits}, along their symbols and symbols referring to the quantity they measure (the ``dimension symbols'').

\begin{table}[h]
  \centering
  $\begin{tabular}{c|c|c|c}
    Unit name & Unit symbol & Dimension symbol & Quantity name\\
	\hline
	second & s & T & time \\
	metre & m & L & length \\
	kilogram & kg & M & mass \\
	ampere & A & I & electric current \\
	kelvin & K & $\Theta$ & temperature \\
	mole & mol & N & amount of substance \\
	candela & cd & J & luminous intensity
  \end{tabular}$
  \caption{SI base units}
  \label{baseUnits}
\end{table}

As of May 2019 all the base units have universal definitions.

It is also important to recognise that the choice of which units are base units is quite arbitrary. We could just as well have chosen speed to be a base units and defined a unit of length based on speed times time. Particularly it might be strange that the ampere is a base units seeing as electric current is electric charge per unit of time. It might seem more intuitive to make coulomb (the SI unit of electric charge) a base unit and define ampere as coulomb per second. The reason this was not done is because historically electric current was easier to measure than electric charge. 

The many other SI units based on these base units will be introduced as necessary in this text.

\section{Geometrized units}
TODO $c = G =1$
Naturalness of numbers and implications for multiplication

TODO See also Mandl, Shaw 6.1

\subsection{Conventions: two fundamental constants}
\begin{align*}
&c = 2,957\times 10^8\text{ms}^{-1} \qquad \to \qquad \text{speed of light}\\
&\hbar \equiv \frac{h}{2\pi} = 6,582\times 10^{-12}\text{MeV s} \qquad \to \qquad \text{Planck constant}
\end{align*}
We set $\hbar = c= 1$ (adimensional), which means we will be working in natural units. This has the advantage of giving simpler expressions, but we lose explicit dimensionality.
\subsection{Assumptions underlying natural units}
\begin{align*}
c =1 \quad &\to \quad [V] =\frac{[L]}{[T]} = \text{dimensionless} \qquad \to \qquad [L] = [T] \\
\hbar = 1 \quad &\to \quad [S] = \frac{[M][L]^{\cancel{2}}}{\cancel{[T]}} = \text{dimemsionless} \qquad \to \qquad [M] = [L]^{-1}
\end{align*}
\[ \Rightarrow [M] = [L]^{-1} = [T]^{-1} \]
\subsection{Dimensions of energy and momentum in natural units}
\begin{align*}
[E] = [M][V]^2 \qquad &\to \qquad [E]=[M] \\
[P] = [M][V] \qquad &\to \qquad [P]=[M]
\end{align*} \[ [M] = c\text{eV} \]


\chapter{Rounding and precision}
significant figures
scientific notation
accuracy vs precision
order of magnitude
dimensions and dimensional analysis (+unit conversion)

\part{Newtonian mechanics}
\setcounter{chapter}{0} % Reset chapter counter
\input{newtonianMechanics}

\part{Analytical mechanics}
\setcounter{chapter}{0} % Reset chapter counter
\input{analyticalMechanics}

\part{Thermodynamics}
\setcounter{chapter}{0} % Reset chapter counter
\input{thermo}
\part{Statistical mechanics}
\setcounter{chapter}{0} % Reset chapter counter
\input{statMech}

\part{Fluid mechanics}
\setcounter{chapter}{0} % Reset chapter counter
\input{fluidMechanics}

\part{Classical electromagnetism}
\setcounter{chapter}{0} % Reset chapter counter
\input{electro}

\part{Optics}
\setcounter{chapter}{0} % Reset chapter counter
\input{optics}


\part{Special relativity}
\setcounter{chapter}{0} % Reset chapter counter
\input{SR}

\part{General relativity}
\setcounter{chapter}{0} % Reset chapter counter
\input{GR}

\part{Quantum mechanics}
\setcounter{chapter}{0} % Reset chapter counter
\input{quantumMechanics}
\input{quantumInformation}
\input{quantum}

\part{Quantum field theory}
\setcounter{chapter}{0} % Reset chapter counter
\input{QFT}

\part{Particle physics}
\setcounter{chapter}{0} % Reset chapter counter
\input{particles}

\part{Quantum many particle systems}
\setcounter{chapter}{0} % Reset chapter counter
\input{quantumManyBody}

\part{Nuclear physics}
\setcounter{chapter}{0} % Reset chapter counter
\input{nuclear}

\part{Solid state physics}
\setcounter{chapter}{0} % Reset chapter counter
\input{solidState}

\part{Soft condensed matter physics and biophysics}
\setcounter{chapter}{0} % Reset chapter counter
\input{softMatter}

\part{Nanophysics}
\setcounter{chapter}{0} % Reset chapter counter
\input{nano}

\part{Plasma physics}
\setcounter{chapter}{0} % Reset chapter counter
\url{http://farside.ph.utexas.edu/teaching/plasma/lectures1/index.html}
    Magneto-hydrodynamics
    Alfvén waves


\part{Cosmology}
\setcounter{chapter}{0} % Reset chapter counter
\input{cosmology}

\part{Reference}
\appendix

\chapter{Symbols}

\chapter{Formula reference}

\chapter{Useful constants}


\chapter{Seminal experiments}
Coulomb torsion balance
Michelson-Morley
Eotvös experiment

Bucket (Newton)
Zero point energy

LIGO
Colliders + CERN

Star composition from spectroscopy

2-degree-Field Galaxy Redshift Survey
2-Micron All-Sky Survey (2MASS)
Bell test
BOOMERanG
Camera obscura experiments
Cavendish experiment
Cosmic Background Explorer (COBE)
Cowan–Reines neutrino experiment
Davisson–Germer
Double-slit (Young)
Foucault pendulum
Franck–Hertz
Gravity Probe A
Gravity Probe B
Geiger–Marsden
Homestake experiment
Michelson–Morley (+ finite $c$?)
Oil drop experiment
Sloan Digital Sky Survey
Stern–Gerlach
Wilkinson Microwave Anisotropy Probe

\chapter{Bibliography}

\url{https://en.wikipedia.org/wiki/Outline_of_physics}

\begin{enumerate}
\item Experimental physics and statistics:
\item Newtonian mechanics:
\begin{itemize}
\item Gregory
\item Goldstein
\item \url{https://en.wikipedia.org/wiki/Principle_of_least_action}
\item The calculus of variations, Bruce van Brunt
\end{itemize}
\item Fluid mechanics:
\item Building blocks of the world:
\item Thermodynamics and statistical mechanics:
\begin{itemize}
\item Schroeder
\item Callen
\item The Concepts and Logic of Classical Thermodynamics as a Theory of Heat Engines: Rigorously Constructed upon the Foundation Laid by S. Carnot and F. Reech; Clifford A. Truesdell, Subramanyam Bharatha 
\item Cengel, Boles
\item Phenomenological thermodynamics in a nutshell, Arnold Neumaier
\item Concepts in thermal physics / Stephen J. Blundell and Katherine M. Blundell
\item Statistical Mechanics, Kerson Huang
\end{itemize}
\item Classical electromagnetism:
\begin{itemize}
\item David J. Griffiths, Introduction to electrodynamics
\end{itemize}
\item Optics:
\begin{itemize}
\item Fundamentals of optics, Jenkins and White
\item The light fantastic, Kenyon
\end{itemize}
\item Special relativity:
\begin{itemize}
\item David J. Griffiths, Introduction to electrodynamics
\item R. D'Auria and M. Trigiante, From special relativity to Feynman diagrams.
\item \url{https://www.quora.com/Which-19th-century-physicist-famously-said-that-all-that-remained-to-be-done-in-physics-was-compute-effects-to-another-decimal-place}
\end{itemize}
\item General relativity:
\begin{itemize}
\item James B. Hartle, Gravity
\item S. Carroll, Spacetime and Geometry: An Introduction to General Relativity. --: Addison-Wesley, 2003
\item M. Blau, Lecture Notes in General Relativity
\item \url{http://casa.colorado.edu/~ajsh/phys5770_08/grtetrad.pdf}
\end{itemize}
\item Quantum mechanics:
\begin{itemize}
\item B.H. Bransden and C.J. Joachain, Quantum mechanics
\item Van Hove, Léon (1958). "Von Neumann's contributions to quantum theory". Bull. Amer. Math. Soc. 64: 95–100. doi:10.1090/s0002-9904-1958-10206-2.
\item SUPPLEMENTARY NOTES ON DIRAC NOTATION,
QUANTUM STATES, ETC.; R. L. Jaffe, 1996
\item A short course on Quantum Mechanics and its Geometry, Elisa Ercolessi
\item Zetilli
\item Townsend
\item Strocchi, F.: An Introduction to the Mathematical Structure of Quantum Mechanics:A Short Course for Mathematicians, vol. 28. World Scientific, New Jersey (2008)
\item Quantum Theory - Peres
\item QM - Schwabl
\item \url{http://www.ueltschi.org/teaching/2022-TCC.html}
\end{itemize}
\item Classical field theory:
\begin{itemize}
\item Nastase, Classical field theory
\end{itemize}
\item Quantum field theory:
\begin{itemize}
\item Perturbative Algebraic Quantum Field Theory, An Introduction for Mathematicians - Kasia Rejzner
\item Klauber, Student friendly QFT
\item Mandl, Shaw
\item Weinberg 
\item Quantum Fields and Strings:  A Course for Mathematicians
\end{itemize}
\item Quantum many particle systems:
\begin{itemize}
\item Quantum theory of many-particle systems, Fetter and Walecka
\end{itemize}
\item Nuclear physics:
\begin{itemize}
\item Introductory nuclear physics, Kenneth Krane
\end{itemize}
\item Solid state physics:
\begin{itemize}
\item Charles Kittel, Introduction to solid state physics
\item John J. Quinn; Kyung-Soo Yi
\item Solid State Physics; Giuseppe Grosso, Giuseppe Pastori Parravicini
\item Ashcroft, Mermin
\item Magnetism and magnetic materials, J.M.D. Coey
\end{itemize}
\item Soft condensed matter:
\begin{itemize}
\item Thermodynamics of the glassy state, Leuzzi and Nieuwenhuizen
\item Kinetics of phase transitions, Puri and Wadhawan (ed.)
\end{itemize}
\item Cosmology:
\end{enumerate}





\end{document}